Master thesis

### 2017-07-13 06-Discussion.Rmd
Discussion v1

**Take home messages:**

## "Model limits"

Discuss limits of the model with sensitivity analysis (insist in seed rain constant importance).

## Diversity and tropical forest ecosystem resilience

* Diversity have a strong role in ecosystem resilience
* Increasing functional diversity results in increased resilience
* Species and functional richnesses is not directly increasing resilience
* But increased species and functional richnesses will increase chance for high functional diversity
* Functional dispersion decreased resilience when functional eveness increased resilience
* Effectively high functional diversity needs eveness in order to better answer disturbance, if not the diversity is masked by ecosystem dominant species

## Biodiversity role in tropical forest ecosystem resilience

* Biodiversity net effect is almost always positive for ecosystems functions
* Biodiversity effects answer to disturbance is split in two parts
* First complementarity effects answer disturbance in first decades whereas the disturbance almost removed selection effect
* Then complementarity effect will diminish toward extinction when selection effect is *prendre le relais* to ensure biodiversity net effect
* This patterns are even stronger for higher disturbances
* In the case of selective logging the complementarity effect will thus be the only one playing between two *rotations*
* High forest diversity with important species complementarity through increased functional diversity is thus an advantage for forest restoration between two rotation in order to maintain both productive ecosystem and sustainable management

## Sylviculture ?

## Selective logging vs. monoculture stand ?

Results advocates for **sustainable** selective logging against monoculture stand: high diversity will increase forest resilience and thus improve logging *turn-overs*. Even if monoculture stand are *irrégularisé* they will not reach mature forest diversity, and consequently they will not reach its natural high resilience. But on the other hand selective logging in tropical forest needs to be sustainable, if not the diversity will slowly decrease after each turn-overs degrading forest ecosystem functions, and resilience will become slower and slower in a *cercle vicieux*. Return toward @Zimmerman2012  discussion.

>>>>

In the limit of the model, we were able to show that diversity improved tropical forest resilience. More particularly, functional diversity and evenness are key components of diversity in forest recovery after disturbance. Moreover, we found that complementarity between species was insuring forest recovery in forest succession start with facilitation before more productive species dominate the forest and insure recovery. Our results advocates for a sustainable harvesting of tropical forests through an increased resilience due to high diversity. But this conclusion should met a sustainable definiton of selective logging following @Zimmerman2012. Because if the harvesting are not sustainable negative feedbaks will slowly diminish diversity and its benefits for forest resilience, resulting in forest degradation.

## TROLL limits

Belowground processes, herbaceaous plants, epyphytes and lianas are not simulated in TROLL but they not reprensent the only limit of the model. Other processes are simulated but simplified, and we used sensitivity analysis to assess their relative importance. We found that few functional traits were influencing whole forest structure and dynamic. In addition, we found that the seed rain constant had an important effect on species functional composition and diversity. High external seed rain resulted in a quick recovery of the system toward an equilibrium close to the regional species frequency levied by the seed rain. On the contrary, low seed rain let the simulated forest works has a closed system with more system feedback but a lower stability through time with longer species diversity transitions. In order to study the role of diversity in forest resilience we decided to remove the seed rain constant to get a closed system and look at the role of diversity when it maintains itself through feedbacks and not with immigration.

## Diversity improve tropical forest resilience

Our results validated the hypothesis of a significative relationship between forest resilience and functional diversity and eveness ($p < 0.01$ and $p < 0.05$ respectivelly). Thus we were able to show that diversity improve forest resilience. More particularly, functional diversity seems a major aspect of resilience if its enforced by a high functional eveness. Effectively high functional diversity needs eveness in order to better answer disturbance, if not the diversity is masked by ecosystem dominant species. Those results confirms the review of @Diaz2001 advocating for underevaluated importance of plant functional diversity in ecosystem processes. Additionally, the role of eveness  confirm the review of @Zhang2012 who highlighted the role of eveness in productivity and thus resilience following our hypothesis. Finally, species and functional richnesses is not directly increasing resilience. But increased species and functional richnesses will increase chance for high functional diversity through the sampling effect [@Loreau1998]. An higher sampling of regional species pool will allow a greater chance to pick more functionally diverse species.

## Complementarity and selection insure forest resilience

We found that complementary effect was insuring forest resilience in the beginning of forest successions. We interpreted this results as the consequence of facilitation processes. As the only ressource simulated by TROLL is the light, the major facilitation will be light shading of post-pioneer species by pioneer species in disturbance gaps. But the complementarity effect is reducing through time, to let the selection effect insure forest resilience due to an inforced dominance of more productive species. The diminishing of complementarity effect is due to competitive selection though time in gaps succession. But the study scale matters as shwon by @Chisholm2013. Here we look at processes to a 16 ha scale. We do not have topography reducing micro-environment effects, but @Chisholm2013 results suggest that complementarity will be stronger at smaller scale and could explain its low value after forest recovery. Finally, our findings confirm results of @Tobner2016 realized on experimental forests of 4 years in low diverse forest of Canada. They also found that selection effect was greater than complementarity in most of the case. In the case of selective logging the complementarity effect will thus be the major effect between two cutting cycles due to the cycle length (several decades in guyana shield). High forest diversity with important species complementarity through increased functional diversity is thus an advantage for forest recovery between two rotation in order to maintain both productive ecosystem and sustainable management.

## Conclusion

We used closed forest system simulated with TROLL forest model to evaluate the role and mechanisms of biodiversity in tropical forest resilience to disturbance. We found that diversity improved forest resilience and that complementarity between species was insuring forest recovery in forest succession start with facilitation before more productive species dominate the forest and insure recovery. Our results advocates for sustainable selective logging against monoculture stand: high diversity will increase forest resilience and thus improve logging cycles. Even if monoculture stand are naturalized they will not reach mature forest diversity, and consequently they will not reach its natural high resilience. But on the other hand selective logging in tropical forest needs to be sustainable, if not the diversity will slowly decrease after each cycles degrading forest ecosystem functions, and resilience will decrease due to negative feedbacks. @Zimmerman2012 criticized the state of sylviculture in tropics, suggesting that "we have not been able to reconcile these opposing biological and economic forces". Still, they advocates for a possibility of sustainable tropical logging, already existing in small-scale, that will need proper funding from the internationnal communtiy. This view correspond to the high resilience, we were able to show, of tropical forest hyperdiverse systems and let hope for a future sustainable selective logging in tropical forests.

### 2017-07-13 docs/master-thesis.tex
Discussion v1

  \begin{center}
    \vspace*{\stretch{1}}
    \LARGE{\textbf{Mémoire de stage}} \\
    \vspace*{\fill}
    \large{présenté par} \\
    \large{Sylvain SCHMITT} \\
    \vspace*{\fill}
    \large{pour obtenir le diplôme national de master} \\
    \large{mention Biodiversité, écologie, évolution} \\
    \small{parcours Biodiversité végétale et gestion des écosystèmes tropicaux (BIOGET)} \\
    \vspace*{\fill}
    \large{Sujet :} \\
    \Large{\textbf{Résilience des écosystèmes forestiers tropicaux après perturbation}} \\
    \vspace*{\fill}
    \large{soutenu publiquement le XX Juin 2017} \\
    \large{à Kourou} \\
    \vspace*{\fill}
    \large{devant le jury suivant :} \\
    \vspace*{\fill}
    Dr Bruno HÉRAULT  \emph{Tuteur de stage} \\
    Titre Prénom NOM  \emph{Examinateur} \\
    Titre Prénom NOM  \emph{Examinateur} \\
    Dr Stéphane TRAISSAC  \emph{Enseignant-réferent} \\
    \vspace*{\stretch{1}}
  \end{center}

@@

\textbf{Take home messages:}

\subsection{\texorpdfstring{``Model
limits''}{Model limits}}\label{model-limits}

Discuss limits of the model with sensitivity analysis (insist in seed
rain constant importance).

\subsection{Diversity and tropical forest ecosystem
resilience}\label{diversity-and-tropical-forest-ecosystem-resilience}

\begin{itemize}
\tightlist
\item
  Diversity have a strong role in ecosystem resilience
\item
  Increasing functional diversity results in increased resilience
\item
  Species and functional richnesses is not directly increasing
  resilience
\item
  But increased species and functional richnesses will increase chance
  for high functional diversity
\item
  Functional dispersion decreased resilience when functional eveness
  increased resilience
\item
  Effectively high functional diversity needs eveness in order to better
  answer disturbance, if not the diversity is masked by ecosystem
  dominant species
\end{itemize}

\subsection{Biodiversity role in tropical forest ecosystem
resilience}\label{biodiversity-role-in-tropical-forest-ecosystem-resilience}

\begin{itemize}
\tightlist
\item
  Biodiversity net effect is almost always positive for ecosystems
  functions
\item
  Biodiversity effects answer to disturbance is split in two parts
\item
  First complementarity effects answer disturbance in first decades
  whereas the disturbance almost removed selection effect
\item
  Then complementarity effect will diminish toward extinction when
  selection effect is \emph{prendre le relais} to ensure biodiversity
  net effect
\item
  This patterns are even stronger for higher disturbances
\item
  In the case of selective logging the complementarity effect will thus
  be the only one playing between two \emph{rotations}
\item
  High forest diversity with important species complementarity through
  increased functional diversity is thus an advantage for forest
  restoration between two rotation in order to maintain both productive
  ecosystem and sustainable management
\end{itemize}

\subsection{Sylviculture ?}\label{sylviculture-1}

\subsection{Selective logging vs.~monoculture stand
?}\label{selective-logging-vs.monoculture-stand}

Results advocates for \textbf{sustainable} selective logging against
monoculture stand: high diversity will increase forest resilience and
thus improve logging \emph{turn-overs}. Even if monoculture stand are
\emph{irrégularisé} they will not reach mature forest diversity, and
consequently they will not reach its natural high resilience. But on the
other hand selective logging in tropical forest needs to be sustainable,
if not the diversity will slowly decrease after each turn-overs
degrading forest ecosystem functions, and resilience will become slower
and slower in a \emph{cercle vicieux}. Return toward
\citet{Zimmerman2012} discussion.

>>>>

  \begin{center}
    \vspace*{\stretch{1}}
    \LARGE{\textbf{Mémoire de stage}} \\
    \vspace*{\fill}
    \large{présenté par} \\
    \large{Sylvain SCHMITT} \\
    \vspace*{\fill}
    \large{pour obtenir le diplôme national de master} \\
    \large{mention Biodiversité, écologie, évolution} \\
    \small{parcours Biodiversité végétale et gestion des écosystèmes tropicaux (BIOGET)} \\
    \vspace*{\fill}
    \large{Sujet :} \\
    \Large{\textbf{Rôle de la biodiversité dans la résilience des écosystèmes forestiers tropicaux après perturbation}} \\
    \vspace*{\fill}
    \large{soutenu publiquement le XX Juin 2017} \\
    \large{à Kourou} \\
    \vspace*{\fill}
    \large{devant le jury suivant :} \\
    \vspace*{\fill}
    Dr Bruno HÉRAULT  \emph{Tuteur de stage} \\
    Titre Prénom NOM  \emph{Examinateur} \\
    Titre Prénom NOM  \emph{Examinateur} \\
    Dr Stéphane TRAISSAC  \emph{Enseignant-réferent} \\
    \vspace*{\stretch{1}}
  \end{center}

@@

In the limit of the model, we were able to show that diversity improved
tropical forest resilience. More particularly, functional diversity and
evenness are key components of diversity in forest recovery after
disturbance. Moreover, we found that complementarity between species was
insuring forest recovery in forest succession start with facilitation
before more productive species dominate the forest and insure recovery.
Our results advocates for a sustainable harvesting of tropical forests
through an increased resilience due to high diversity. But this
conclusion should met a sustainable definiton of selective logging
following \citet{Zimmerman2012}. Because if the harvesting are not
sustainable negative feedbaks will slowly diminish diversity and its
benefits for forest resilience, resulting in forest degradation.

\subsection{TROLL limits}\label{troll-limits}

Belowground processes, herbaceaous plants, epyphytes and lianas are not
simulated in TROLL but they not reprensent the only limit of the model.
Other processes are simulated but simplified, and we used sensitivity
analysis to assess their relative importance. We found that few
functional traits were influencing whole forest structure and dynamic.
In addition, we found that the seed rain constant had an important
effect on species functional composition and diversity. High external
seed rain resulted in a quick recovery of the system toward an
equilibrium close to the regional species frequency levied by the seed
rain. On the contrary, low seed rain let the simulated forest works has
a closed system with more system feedback but a lower stability through
time with longer species diversity transitions. In order to study the
role of diversity in forest resilience we decided to remove the seed
rain constant to get a closed system and look at the role of diversity
when it maintains itself through feedbacks and not with immigration.

\subsection{Diversity improve tropical forest
resilience}\label{diversity-improve-tropical-forest-resilience}

Our results validated the hypothesis of a significative relationship
between forest resilience and functional diversity and eveness
(\(p < 0.01\) and \(p < 0.05\) respectivelly). Thus we were able to show
that diversity improve forest resilience. More particularly, functional
diversity seems a major aspect of resilience if its enforced by a high
functional eveness. Effectively high functional diversity needs eveness
in order to better answer disturbance, if not the diversity is masked by
ecosystem dominant species. Those results confirms the review of
\citet{Diaz2001} advocating for underevaluated importance of plant
functional diversity in ecosystem processes. Additionally, the role of
eveness confirm the review of \citet{Zhang2012} who highlighted the role
of eveness in productivity and thus resilience following our hypothesis.
Finally, species and functional richnesses is not directly increasing
resilience. But increased species and functional richnesses will
increase chance for high functional diversity through the sampling
effect \citep{Loreau1998}. An higher sampling of regional species pool
will allow a greater chance to pick more functionally diverse species.

\subsection{Complementarity and selection insure forest
resilience}\label{complementarity-and-selection-insure-forest-resilience}

We found that complementary effect was insuring forest resilience in the
beginning of forest successions. We interpreted this results as the
consequence of facilitation processes. As the only ressource simulated
by TROLL is the light, the major facilitation will be light shading of
post-pioneer species by pioneer species in disturbance gaps. But the
complementarity effect is reducing through time, to let the selection
effect insure forest resilience due to an inforced dominance of more
productive species. The diminishing of complementarity effect is due to
competitive selection though time in gaps succession. But the study
scale matters as shwon by \citet{Chisholm2013}. Here we look at
processes to a 16 ha scale. We do not have topography reducing
micro-environment effects, but \citet{Chisholm2013} results suggest that
complementarity will be stronger at smaller scale and could explain its
low value after forest recovery. Finally, our findings confirm results
of \citet{Tobner2016} realized on experimental forests of 4 years in low
diverse forest of Canada. They also found that selection effect was
greater than complementarity in most of the case. In the case of
selective logging the complementarity effect will thus be the major
effect between two cutting cycles due to the cycle length (several
decades in guyana shield). High forest diversity with important species
complementarity through increased functional diversity is thus an
advantage for forest recovery between two rotation in order to maintain
both productive ecosystem and sustainable management.

\subsection{Conclusion}\label{conclusion}

We used closed forest system simulated with TROLL forest model to
evaluate the role and mechanisms of biodiversity in tropical forest
resilience to disturbance. We found that diversity improved forest
resilience and that complementarity between species was insuring forest
recovery in forest succession start with facilitation before more
productive species dominate the forest and insure recovery. Our results
advocates for sustainable selective logging against monoculture stand:
high diversity will increase forest resilience and thus improve logging
cycles. Even if monoculture stand are naturalized they will not reach
mature forest diversity, and consequently they will not reach its
natural high resilience. But on the other hand selective logging in
tropical forest needs to be sustainable, if not the diversity will
slowly decrease after each cycles degrading forest ecosystem functions,
and resilience will decrease due to negative feedbacks.
\citet{Zimmerman2012} criticized the state of sylviculture in tropics,
suggesting that ``we have not been able to reconcile these opposing
biological and economic forces''. Still, they advocates for a
possibility of sustainable tropical logging, already existing in
small-scale, that will need proper funding from the internationnal
communtiy. This view correspond to the high resilience, we were able to
show, of tropical forest hyperdiverse systems and let hope for a future
sustainable selective logging in tropical forests.
